Tail Flexibility in the Degrees of Preferential Attachment Networks

### 2025-05-08 paper.tex
Add temporary section of comments w/ some minor edits in qmd

% author setup
\author[1]{\fnm{Thomas William} \sur{Boughen}}\author[1]{\fnm{Clement} \sur{Lee}}\author[1]{\fnm{Vianey Palacios} \sur{Ramirez}}
% affil setup
\affil[1]{\orgdiv{School of Mathematics, Statistics and
Physics}, \orgname{Newcastle University}}

@@

@@

If the power law is an adequate model to use for real degree
distributions has become a heated debate, with alternatives such as
\citet{Broido_2019} seeming to outperform the power law in some cases.
Additionally, \citet{lef24} find that for many real degree distributions
the body may follow a power law, the right tail does not or at least
seems to be lighter than is implied by the body whilst still being
heavy. One method for modelling the tail of the degree distributions of
real networks is to use a discretised variation of the generalised
Pareto distribution (GPD) usually dubbed the Integer GPD (GPD) where
\(X|X>v \sim \mathrm{IGPD}(\xi,\sigma,v)\) is defined by the survival:

@@

\citep[CL: Is the precise form of the IGPD needed here? The big picture
is probably more important: extreme value methods have not been used in
analysing degree distributions until recently, such as][ and
\citet{lef24}. Not using these methods essentially downweights the
deviation of the vertices with large degrees, which have far more
influence on the network than those with small degrees, from the power
law behaviour.]{vvvk19}

@@

Estimating the preference function used by network generative models
like this is of great interest, with the potential to provide a deeper
understanding of the relationships between the vertices in the network.
Several papers and at least one R package (\texttt{PAfit}) have been
dedicated to estimating the preference function that may have been used
when a network was growing; one thing all these have in common is
requiring access to a networks evolution data, something that is not
always readily available or feasible to access. A method for estimating
a preference function without this data available would be extremely
useful.

\citep{rudas07} introduces theoretical results for this model with a
general preference function in the cases that the network being
generated is a tree (\(m=1\)); presenting an opportunity to design a
preference function that will capture the behaviours observed in the
degree distributions of real networks. Since they provide theoretical
results for the limiting degree distribution, a model of this form would
be able to be fitted to the degree distributions of real networks where
the full evolution is not available, only a snapshot.

The rest of this paper is as follows: Section~\ref{sec-gpa} looks at the
theoretical limiting behaviour of a GPA model when \(m=1\) according to
the results from \citep{rudas07}, going on to introduce a flexible class
of preference functions that can guarantee a heavy tailed degree
distribution while still being flexible in the body.
Section~\ref{sec-rec} shows that the preference function parameters can
be fairly well recovered from the degree distributions of networks
simulated from the GPA model using various functions of the class
introduced in Section~\ref{sec-gpa}. Having shown that for simulated
data the model parameters are recoverable from the degree distribution
alone, Section~\ref{sec-real} builds upon this and attempts to fit the
model to degree distributions of real networks; modelling not only the
degree distribution itself but also the estimates for the parameters of
the preference function assuming the networks evolved according to the
GPA scheme. Ultimately, Section~\ref{sec-conc} discusses the main
results of the paper, pitfalls of this method and future work.

\section{General Preferential Attachment}\label{sec-gpa}

The focus of this section is the limiting degree distribution of a
preferential attachment network with preference function \(b(\cdot)\)
where each vertex brings with it \(m\) edges. This preference function
is subject to the following conditions:

@@

Given these two conditions an expression for the survival of the
limiting degree distribution can be found in the case that \(m=1\);
obtained by considering a branching process that is equivalent to the
growth of the network, as in \citep{rudas07}. Theorem 1 from
\citep{rudas07} states that for the tree \(\Upsilon(t)\) at time \(t\):

\[
\lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda^* \int_0^\infty e^{-\lambda^* t}\mathbb E\left[\varphi(\Upsilon(t))\right]dt
\] where \(\lambda^*\) satisfies \(\hat\rho(\lambda^*)=1\).

@@

\[
\bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}
\] Additionally, using the fact that \(f(k) = \bar F(k-1) - \bar F(k)\)
the probability mass function (p.m.f.) can be shown to be:

\[
f(k) = \frac{\lambda^*}{\lambda^* + b(k)}\prod_{i=0}^{k-1}\frac{b(i)}{\lambda^*+b(i)},\qquad k\in\mathbb N.
\]

@@

\[
\lim_{k\rightarrow\infty}\Omega(F,n) = \lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*}.
\]

@@

Previously, in \citet{krapivsky01}, a preference function of the form
\(k^\alpha +\varepsilon\) has been considered. However, it has been
noted that super-linear preferential attachment (\(\alpha > 1\)) leads
to a degenerate limiting degree distribution where at some point one
vertex gains the connection from every vertex that joins the network,
additionally sub-linear preferential attachment (\(\alpha <1\)) has been
shown to lead to a light tailed limiting degree distribution. Consider a
preference function of the form:

\[
b(k) = \begin{cases}
k^\alpha + \varepsilon,&k<k_0\\
k_0^\alpha + \varepsilon + \beta(k-k_0), &k\ge k_0
\end{cases}
\] for \(\alpha,\beta, \varepsilon>0\) and \(k_0\in\mathbb N\).

@@

\begin{equation}\phantomsection\label{eq-polysurv}{
\bar F(k) = \begin{cases}
\displaystyle\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0\\
\displaystyle\left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right)\frac{\Gamma(\lambda^*+k_0^\alpha + \varepsilon)/\beta)}{\Gamma\left((k_0^\alpha + \varepsilon)/\beta\right)} \frac{\Gamma\left(k-k_0 + 1 +\frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0 + 1 +\frac{\lambda^* +k_0^\alpha + \varepsilon}{\beta}\right)},&k\ge k_0,
\end{cases}
}\end{equation}

with \(\lambda^*\) satisfying

\[
\hat\rho(\lambda^{*}) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda^{*}+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda^{*}-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^{*} + i^\alpha + \varepsilon} = 1,
\]

which has to be solved numerically for most parameter choices. Note that
\(\lambda>\beta\) since when \(\lambda\le \beta\) the infinite sum no
longer converges and instead goes to infinity.

@@

\[
\bar F(k) 
%\begin{cases}
%=\displaystyle\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0,\\
\displaystyle\approx \left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right) \left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^*/\beta},%&k\ge k_0,
%\end{cases}
\] which is the survival function of the
\(\text{IGPD}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)\)
distribution. This means that, above the threshold of the piecewise
preference function, the limiting degree distribution can be
approximated by the IGPD with shape parameter \(\beta/\lambda^{*}\).

@@

Some examples of the possible degree distributions are shown in the
central row of Figure~\ref{fig-polylinsurv}.

@@

The procedure for recovering the parameters begins with simulating a
network from the model with \(N=100,000\) vertices and \(m=1\) given
some set of parameters
\(\pmb\theta = (\alpha, \beta, \varepsilon, k_0)\), obtaining the degree
counts in the form of a vector of degrees \(\pmb x = (1,2,\ldots,M)\)
and the number of vertices with those degrees \(\pmb n\) where \(M\) is
the maximum degree. However, when it comes to fitting this model to real
data in the next section, small degrees are extremely influential and
will often cause the model to have a bad fit even though the rest of the
data looks similar to that of the simulations; truncating the data above
some value, say \(l<k_0\), should allow the model to be more effectively
fitted. This accounts in some way for the fact that this model only
applies to trees, and could not possibly capture the behaviour for nodes
with small degrees. Using the p.m.f derived in \citep{rudas07}, the
likelihood given that \(x_i \ge l\) for all \(i =1,2,\ldots,M\) is:

@@

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-rec1-1.pdf}

@@

\caption{\label{fig-rec1}Posterior estimates of survival function for
data simulated from the proposed model with various combinations of
(\(\alpha\),\(\beta\),\(\varepsilon\)) and \(k_0=20\).}

>>>>

% author setup
\author[1]{\fnm{Thomas William} \sur{Boughen}}\author[1]{\fnm{Clement} \sur{Lee}}\author[1]{\fnm{Vianey Palacios} \sur{Ramirez}}
% affil setup
\affil[1]{\orgdiv{School of Mathematics, Statistics and
Physics}, \orgname{Newcastle University, UK}}

@@

\section*{CL: Overall comments}\label{cl-overall-comments}
\addcontentsline{toc}{section}{CL: Overall comments}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The introduction section is not quite clear and needs a lot of
  improvement, so here are my suggestions on the outline, in a hopefully
  more logical flow:

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    Start with the broad picture e.g.~that networks exist in many fields
    (name some for which you can cite applications), and that
    statistical methods are have been used to study networks. Some are
    for revealing the latent / underlying structures of the networks
    (stochastic block models), some are for explaining some response
    variables by network properties (ERGMs), and some are for
    postulating how the network grows and evolves (mechanistic models).
    Find some references, either the groundbreaking ones or
    comprehensive reviews to back this up.
  \item
    Introduce the general PA model as one of the mechanistic models.
    This will be a natural point to mention the rules (in terms as
    simple as possible) and introduce the preference function. Mention
    the special case when \(b(k)=k\) and cite \citet{Barabasi99}.
  \item
    Then talk about PA model and mechanistic models in general being
    useful to \textbf{simulate} networks but not be fitted to real data
    because usually only snapshots are available. This leads well into
    the point on the degree distribution being informative.
  \item
    Then in the paragraph starting ``When looking at \ldots{}'', point
    out that it is the BA model \citep{Barabasi99} that leads to a power
    law degree distribution.
  \item
    Then point out that while it's attempting to attribute to the BA
    model when degrees seemingly follow the power law, there are two
    main concerns: first, a general preference function doesn't lead to
    the power law \citep{krapivsky01}; second, some claim that a lot of
    real data don't actually follow the power law. Mention not only
    \citet{Broido_2019} but also responses to them e.g. \citet{vvvk19}.
  \item
    \citet{vvvk19} leads naturally to the point that extreme value
    methods have not been used in analysing degree distributions until
    recently, such as \citet{vvvk19} and \citet{lef24}. Not using these
    methods essentially downweights the deviation of the vertices with
    large degrees, which have far more influence on the network than
    those with small degrees, from the power law behaviour.
  \item
    Also mention Resnick's works (and others) on the crossroads of
    networks and extremes. Summarise what these papers do.
  \item
    Come back to the focus on the degree distribution, and point out
    that most methods, whether extremes are incorporated or not, are
    descriptive in the sense that no information about the preference
    function \(b(\cdot)\) is revealed, even if PA model was indeed the
    underlying mechanism.
  \item
    Then mention that this paper addresses that gap. Specifically,
    \textbf{given the degree distribution of the network assumed to come
    from the PA model, can we directly infer the model parameters?}
    While \citet{rudas07} have derived the limiting degree distribution
    in terms of \(b(\cdot)\) and its parameters, we build upon it by
    finding the class of preference functions that result in realistic
    tails (heavy but not as heavy as implied by the power law).
  \item
    Then go straight to the division of the rest of the paper, and
    conclude the introduction section. The remaining materials should
    mostly go to the next section, but you can retain some materials if
    you think they complement (or already address) the above points.
  \end{enumerate}
\item
  Rough outline of Section 2, which should be on the PA model and
  deriving the tail heaviness of the limiting degree distribution in
  terms of \(b(\cdot)\):

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    Start with the general description of the PA model. You can move
    here the paragraphs in the introduction section, starting with ``The
    network generative model \ldots{}'', ``Starting at time \ldots{}'',
    and ``Special cases of this model \ldots{}''. Most of what you have
    written up till the pmf can be kept.
  \item
    Next, first state that of interest is the tail heaviness (which here
    refers to the shape parameter of the (integer) GPD approximation of
    the tail). Then move here the paragraphs on \citet{shimura12} from
    the introduction.
  \item
    Then state proposition 2.1 and point to the Appendix for the proof.
  \item
    Next, we need a bit more discussion and explanation before stating
    that \(b(\cdot)\) has to be asymptotically linear for a heavy tail.
    First point out that this result covers and aligns with the
    previously obtained result i.e. \citet{krapivsky01}. This is pretty
    much the paragraph currently under Section 2.1, but with a better
    flow. You might even include the special case of the special case
    i.e.~\(b(k)=k+\epsilon\), to briefly illustrate how a tail heaviness
    of \(0.5\) is obtained for the BA model.
  \item
    Now, argue that \(b(\cdot)\) has to be (asymptotically) linear above
    a threshold, in order to have a positive tail heaviness, then
    introduce the proposed preference function as our model, with tail
    heaviness \(\beta/\lambda^{*}\).
  \item
    Conclude the section by saying that we will examine its property in
    the next Section (see next point).
  \end{enumerate}
\item
  The section could be called ``Preferential attachment with flexible
  heavy tail'', essentially making the subsection a section itself.

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    Most of the materials can be kept, and the form of the IGPD should
    be inserted when approximating the survival using Stirling's (not
    Sterling's) approximation. The survival of the IGPD was in the
    introduction but I suggest moving it here.
  \item
    The piecewise linear model doesn't add much, so I would consider
    excluding it unless there's a point e.g.~computational simplicity?
  \item
    I would conclude the section by moving the likelihood expression
    here, and say that the parameters can be inferred in a standard way,
    via the frequentist or Bayesian approach, as the likelihood is an
    explicit function of the parameters.
  \end{enumerate}
\item
  For simplicity, it's easier to use \(m=1\) throughout and only mention
  the general case \(m>1\) as future work in the discussion section.
  Also, it might be worth just calling the general preferential
  attachment model without the word ``general'' throughout, as it is
  unambiguous that it is the model with a flexible / general
  \(b(\cdot)\), while the BA model is the one with \(\b(k)=k+\epsilon\).
\item
  There are inconsistencies and duplications in various places, probably
  because you edited at one place and didn't update associated instances
  elsewhere. Do go back and forth when making edits.
\item
  The presentation can use many incremental improvements:

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    use \(k\) or \(n\) for the integer but not both, unless necessary
    and in a consistent way. There were places where \(n\) was used as
    the subject initially but it changed to \(k\).
  \item
    ensure that the text in plots are legible.
  \item
    when citing, use the parentheses correctly, as they have been used
    inappropriately; in Quarto syntax, the square brackets will
    automatically put the parentheses (similar to citep in LaTeX).
  \item
    ensure that the maths don't over the linewidth.
  \item
    introduce the full term with the acronym in parentheses when using
    it for the first time, e.g.~GPA.
  \item
    some references of equations are not working e.g.~eq-surv?
  \item
    unless necessary, avoid strong adverbs / adjectives.
  \end{enumerate}
\item
  It seems that you version control all the files, which I won't
  recommend. Only track the files that you manually change, not the
  generated ones (cache, pdf, and tex unless you manually created one).
  This will facilitate the reproducibility. I have removed the generated
  files in my branch.
\item
  Consider creating subfolders to store the results, data, figures, etc.
  This will streamline your file structure. I have accordingly edited
  the R scripts so that if you run them, the rds files will be stored in
  results/.
\item
  Accents in names in bib / tex / qmd might not work on some computers.
  Here's the conversion chart: https://www.bibtex.org/SpecialSymbols/
\end{enumerate}

@@

Whether the power law is adequate for real degree distributions has
become a heated debate, with alternatives such as \citet{Broido_2019}
seeming to outperform the power law in some cases. Additionally,
\citet{lef24} find that for many real degree distributions the body may
follow a power law, the right tail does not or at least seems to be
lighter than is implied by the body whilst still being heavy. One method
for modelling the tail of the degree distributions of real networks is
to use a discretised variation of the generalised Pareto distribution
(GPD) usually dubbed the Integer GPD (GPD) where
\(X|X>v \sim \mathrm{IGPD}(\xi,\sigma,v)\) is defined by the survival:

@@

{[}CL: Is the precise form of the IGPD needed here? The big picture is
probably more important (point 3f in overall comments){]}

@@

Estimating the preference function used by network generative models
like this is of great interest, with the potential to provide a deeper
understanding of the relationships between the vertices in the network.
Several papers and an R package (\texttt{PAfit}) have been dedicated to
estimating the preference function that may have been used when a
network was growing; one thing all these have in common is requiring
access to a networks evolution data, something that is not always
readily available or feasible to access. A method for estimating a
preference function without this data available would be extremely
useful.

\citet{rudas07} introduces theoretical results for this model with a
general preference function in the cases that the network being
generated is a tree (\(m=1\)); presenting an opportunity to design a
preference function that will capture the behaviours observed in the
degree distributions of real networks. Since they provide theoretical
results for the limiting degree distribution, a model of this form would
be able to be fitted to the degree distributions of real networks where
the full evolution is not available, only a snapshot.

The rest of this paper is as follows: Section~\ref{sec-gpa} looks at the
theoretical limiting behaviour of a GPA model when \(m=1\) according to
the results from \citet{rudas07}, going on to introduce a flexible class
of preference functions that can guarantee a heavy tailed degree
distribution while still being flexible in the body.
Section~\ref{sec-rec} shows that the preference function parameters can
be fairly well recovered from the degree distributions of networks
simulated from the GPA model using various functions of the class
introduced in Section~\ref{sec-gpa}. Having shown that for simulated
data the model parameters are recoverable from the degree distribution
alone, Section~\ref{sec-real} builds upon this and attempts to fit the
model to degree distributions of real networks; modelling not only the
degree distribution itself but also the estimates for the parameters of
the preference function assuming the networks evolved according to the
GPA scheme. Section~\ref{sec-conc} discusses the main results, proposes
future work and concludes the paper.

\section{General Preferential Attachment Model}\label{sec-gpa}

In this section, we introduce a general preferential attachment model,
with the focus on the limiting degree distribution as a function of the
preference function \(b(\cdot)\) where each vertex brings with it \(m\)
edges. This preference function is subject to the following conditions:

@@

Given these two conditions an expression for the survival of the
limiting degree distribution can be found in the case that \(m=1\);
obtained by considering a branching process that is equivalent to the
growth of the network, as in \citet{rudas07}. Theorem 1 from
\citet{rudas07} states that for the tree \(\Upsilon(t)\) at time \(t\):

\[
\lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda^* \int_0^\infty e^{-\lambda^* t}\mathbb E\left[\varphi(\Upsilon(t))\right]dt
\] where \(\lambda^*\) satisfies \(\hat\rho(\lambda^*)=1\). (CL:
\(\hat{\rho}\) has not been defined yet.)

@@

\[
\bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}.
\] Using the fact that \(f(k) = \bar F(k-1) - \bar F(k)\) the
probability mass function (pmf) is

\[
f(k) = \frac{\lambda^*}{\lambda^* + b(k)}\prod_{i=0}^{k-1}\frac{b(i)}{\lambda^*+b(i)},\qquad k\in\mathbb N,
\]

@@

\[
\lim_{k\rightarrow\infty}\Omega(F,k) = \lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*}.
\]

@@

In \citet{krapivsky01}, a preference function of the form
\(k^\alpha +\varepsilon\) has been considered. However, it has been
noted that super-linear preferential attachment (\(\alpha > 1\)) leads
to a degenerate limiting degree distribution where at some point one
vertex gains the connection (CL: edge?) from every vertex that joins the
network, additionally sub-linear preferential attachment (\(\alpha <1\))
has been shown to lead to a light tailed limiting degree distribution.
Consider a preference function of the form:

\[
b(k) = \begin{cases}
k^\alpha + \varepsilon,&k<k_0,\\
k_0^\alpha + \varepsilon + \beta(k-k_0), &k\ge k_0,
\end{cases}
\] for \(\alpha,\beta, \varepsilon>0\) and \(k_0\in\mathbb N\).

@@

\begin{equation}\phantomsection\label{eq-polysurv}{
\bar F(k) = \begin{cases}
\displaystyle\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0,\\
\displaystyle\left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right)\frac{\Gamma(\lambda^*+k_0^\alpha + \varepsilon)/\beta)}{\Gamma\left((k_0^\alpha + \varepsilon)/\beta\right)} \frac{\Gamma\left(k-k_0 + 1 +\frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0 + 1 +\frac{\lambda^* +k_0^\alpha + \varepsilon}{\beta}\right)},&k\ge k_0,
\end{cases}
}\end{equation}

with \(\lambda^*\) satisfying \(\hat{\rho}(\lambda^{*})=1\), where

\[
\hat\rho(\lambda) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda + i^\alpha + \varepsilon} = 1,
\]

which has to be solved numerically for most parameter choices. Note that
\(\lambda\) has to be greater than \(\beta\) since when
\(\lambda\le \beta\) the infinite sum diverges. (CL: Where is the
infinite sum?)

@@

\[
\bar F(k) 
%\begin{cases}
%=\displaystyle\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0,\\
\displaystyle\approx \left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right) \left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^*/\beta},%&k\ge k_0,
%\end{cases}
\] which is the survival function of the
\(\text{IGPD}\displaystyle\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)\)
distribution. This means that, above the threshold of the piecewise
preference function, the limiting degree distribution can be
approximated by the IGPD with shape parameter \(\beta/\lambda^{*}\).

@@

Some examples of the limiting degree distributions are shown in the
central column of Figure~\ref{fig-polylinsurv}.

@@

The procedure for recovering the parameters begins with simulating a
network from the model with \(N=100,000\) vertices and \(m=1\) given
some set of parameters
\(\pmb\theta = (\alpha, \beta, \varepsilon, k_0)\), obtaining the degree
counts in the form of a vector of degrees \(\pmb x = (1,2,\ldots,M)\)
and the number of vertices with those degrees \(\pmb n\) where \(M\) is
the maximum degree. However, when it comes to fitting this model to real
data in the next section, small degrees are extremely influential and
will often cause the model to have a bad fit even though the rest of the
data looks similar to that of the simulations; truncating the data above
some value, say \(l<k_0\), should allow the model to be more effectively
fitted. This accounts in some way for the fact that this model only
applies to trees, and could not possibly capture the behaviour for nodes
with small degrees. Using the p.m.f derived in \citet{rudas07}, the
likelihood given that \(x_i \ge l\) for all \(i =1,2,\ldots,M\) is:

@@

\includegraphics[width=0.95\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-rec1-1.pdf}

@@

\caption{\label{fig-rec1}Posterior estimates of survival function for
data simulated from the proposed model with various combinations of
(\(\alpha,\beta,\varepsilon\)) and \(k_0=20\).}
